---
title: "Models Available in DSL"
output: 
  md_document:
    variant: markdown_github
---

<div class="card bg-light mb-3">
<div class="card-header">**Table of Contents**</div>
<div class="card-body">

- [**`lm`: Linear Regression**](#lm) <br>

- [**`logit`: Logistic Regression**](#logit) <br>

- [**`poisson`: Poisson Regression**](#poisson) <br>

- [**`felm`: Fixed Effects Linear Regression**](#felm) <br>

- [**`fepois`: Fixed Effects Poisson Regression**](#fepois) <br>

- [**`negbin`: Negative Binomial Regression**](#negbin) <br>

- [**`fenegbin`: Fixed Effects Negative Binomial Regression**](#fenegbin) <br>


</div>
</div>

Note: We currently cover seven models (`lm`, `logit`, `poisson`, `felm`, `fepois`, `negbin`, `fenegbin`). In the next update, we are planning to add more models, including multinomial-logit regression and two-stage least squares for the instrumental variable method. 

<br>

### **`lm`: Linear Regression**{#lm} 
<p class="lh-lg">
</p>

Researchers can implement DSL linear regression by specifying `model = "lm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
library(dsl)
data("data_lm") # example data 
head(data_lm)
```

In this example, variable `Y` requires text annotation (they take `NA` if a given observation is not labeled by experts). Variable `pred_Y` represents predictions for `Y`.

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_lm <- dsl(model = "lm", 
              formula = Y ~ X1 + X2 + X3 + X4 + X5,
              predicted_var = "Y",
              prediction = "pred_Y",
              data = data_lm)
```

| Argument | Description              |
|:-------- | :------------------------|
|`model` | A regression model. `dsl` currently supports `lm` (linear regression), `logit` (logistic regression), `poisson` (Poisson regression), `felm` (fixed-effects regression), `fepois` (fixed-effects Poisson regression), `negbin` (negative binomial regression), and `fenegbin` (fixed-effects negative binomial regression).|
|`formula` | A formula used in the specified regression model.|
|`predicted_var` | A vector of column names in the data that correspond to variables that need to be predicted.|
|`prediction` | A vector of column names in the data that correspond to predictions of `predicted_var`. |
| `data`    | A data frame. The class should be `data.frame`.|

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
summary(out_lm)
```

<br>

### **`logit`: Logistic Regression**{#logit} 
<p class="lh-lg">
</p>

Researchers can implement DSL logistic regression by specifying `model = "logit"`. It requires the same arguments as `model = "lm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_logit") # example data 
head(data_logit)
```

In this example, variables `Y` and `X1` require text annotation (they take `NA` if a given observation is not labeled by experts). Variables `pred_Y` and `pred_X1` represent predictions for `Y` and `X1`. As in this example, `dsl` can handle cases when multiple variables require text annotations.

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_logit <- dsl(model = "logit", 
                 formula = Y ~ X1 + X2 + X4,
                 predicted_var = c("Y", "X1"),
                 prediction = c("pred_Y", "pred_X1"),
                 data = data_logit)
summary(out_logit)
```

<br>

### **`poisson`: Poisson Regression**{#poisson} 
<p class="lh-lg">
</p>

Researchers can implement DSL Poisson regression for count outcomes by specifying `model = "poisson"`. It requires the same arguments as `model = "lm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_poisson") # example data 
head(data_poisson)
```

In this example, variable `Y` is a count (e.g., the number of times a topic is mentioned in a document) that requires text annotation (it takes `NA` if a given observation is not labeled by experts). Variable `pred_Y` represents predicted counts for `Y`.

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_poisson <- dsl(model = "poisson", 
                   formula = Y ~ X1 + X2 + X3 + X4,
                   predicted_var = "Y",
                   prediction = "pred_Y",
                   data = data_poisson)
summary(out_poisson)
```

As in a standard Poisson regression, coefficients are on the log scale, and `exp()` of a coefficient is the multiplicative change in the expected count (the incidence rate ratio). Because `dsl` estimates standard errors with the robust (sandwich) variance, `model = "poisson"` is Poisson Pseudo Maximum Likelihood (PPML). It only assumes that the expected count is `exp()` of the linear predictor, so the outcome does not need to follow a Poisson distribution (e.g., overdispersed counts are allowed) or be an integer. It should be non-negative.

<br>

### **`felm`: Fixed Effects Linear Regression**{#felm} 
<p class="lh-lg">
</p>

Researchers can implement DSL fixed effects linear regression by specifying `model = "felm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_lm")  # example data 
head(data_felm)
```

In this example, variable `log_gsp` requires text annotation (it takes `NA` if a given observation is not labeled by experts). Variable `pred_log_gsp` represents predictions for `log_gsp`. 


#### One-way Fixed Effects

To implement DSL one-way fixed effects regression, users can set `fixed_effect = "oneway"`. Use `index` to denote which column defines fixed effects. Users can also cluster standard errors by specifying a variable name in argument `cluster` (this `cluster` argument is also avaiable for other models). 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_felm_one <- dsl(model = "felm", 
                    formula = log_pcap ~ log_gsp + log_pc + unemp,
                    predicted_var =  "log_gsp",
                    prediction = "pred_log_gsp",
                    fixed_effect = "oneway", 
                    index = c("state"), 
                    cluster = "state",
                    data = data_felm)
```
| Argument | Description              |
|:-------- | :------------------------|
|`fixed_effect` | A type of fixed effects regression you run. `oneway` (one-way fixed effects) or `twoways` (two-way fixed effects).|
|`index` | A vector of column names specifying fixed effects. When `fixed_effect = oneway`, it has one element. When `fixed_effect = twoways`, it has two elements, e.g., `index = c("state", "year")`. |
|`cluster` | A column name in the data that indicates the level at which cluster standard errors are calculated. Default is `NULL`.|


```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
summary(out_felm_one)
```

#### Two-way Fixed Effects

To implement DSL two-way fixed effects regression, users can set `fixed_effect = "twoways"`. Use `index` (a vector of length 2) to denote which columns define fixed effects. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_felm_two <- dsl(model = "felm", 
                    formula = log_pcap ~ log_gsp + log_pc + unemp,
                    predicted_var =  "log_gsp",
                    prediction = "pred_log_gsp",
                    fixed_effect = "twoways", 
                    index = c("state", "year"), 
                    cluster = "state",
                    data = data_felm)
```

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
summary(out_felm_two)
```

<br>

### **`fepois`: Fixed Effects Poisson Regression**{#fepois} 
<p class="lh-lg">
</p>

Researchers can implement DSL fixed effects Poisson regression for count outcomes by specifying `model = "fepois"`. It uses the same arguments as `model = "felm"` (`fixed_effect`, `index`, and `cluster`). 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_fepois") # example data 
head(data_fepois)
```

In this example, variable `Y` is a count that requires text annotation (it takes `NA` if a given observation is not labeled by experts), and variable `pred_Y` represents predicted counts for `Y`. The data is a panel of 100 units over 20 years (`unit` and `year`). 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_fepois <- dsl(model = "fepois", 
                  formula = Y ~ X1 + X2,
                  predicted_var = "Y",
                  prediction = "pred_Y",
                  fixed_effect = "twoways", 
                  index = c("unit", "year"), 
                  cluster = "unit",
                  data = data_fepois)
summary(out_fepois)
```

As in `model = "poisson"`, coefficients are on the log scale and the estimator is Poisson Pseudo Maximum Likelihood (PPML), which allows overdispersed and non-integer outcomes. `fixed_effect = "oneway"` with one element in `index` implements one-way fixed effects.

We recommend `model = "fepois"` over adding fixed effects as dummy variables (e.g., `factor(unit)`) to `model = "poisson"`. The DSL estimator of the fixed effect of each group relies on the DSL-corrected sum of the outcome within the group. When a group is small, one labeled observation with a large prediction error can make this sum negative, and the fixed effect (the log of a negative number) is not defined. `model = "fepois"` estimates the fixed effects on the exponentiated scale, where they are defined even in such cases, and concentrates out the fixed effects of the index with more levels (with `fixed_effect = "twoways"`, the fixed effects of the other index are estimated together with the coefficients). When all fixed effects are defined, it gives the same estimates and standard errors as the dummy-variable approach. 

<br>

### **`negbin`: Negative Binomial Regression**{#negbin} 
<p class="lh-lg">
</p>

Researchers can implement DSL negative binomial regression for overdispersed count outcomes (the variance is larger than the mean) by specifying `model = "negbin"`. It requires the same arguments as `model = "poisson"`. The outcome in `data_poisson` is overdispersed.

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_negbin <- dsl(model = "negbin", 
                  formula = Y ~ X1 + X2 + X3 + X4,
                  predicted_var = "Y",
                  prediction = "pred_Y",
                  data = data_poisson)
summary(out_negbin)
```

The negative binomial model assumes `Var(Y | X) = mu + mu^2/theta`, where `theta` is the dispersion parameter (reported at the bottom of `summary()` and stored in `out_negbin$theta`). Coefficients are on the log scale, as in `model = "poisson"`. They are estimated by maximum likelihood, and `theta` is estimated by maximum likelihood using the labeled observations (predictions carry no information about the dispersion of the outcome). Because standard errors are robust (sandwich), the coefficients and their standard errors remain valid when the expected count is correctly specified, even if the outcome does not follow the negative binomial distribution. 


<br>

### **`fenegbin`: Fixed Effects Negative Binomial Regression**{#fenegbin} 
<p class="lh-lg">
</p>

Researchers can implement DSL fixed effects negative binomial regression by specifying `model = "fenegbin"`. It uses the same arguments as `model = "fepois"` (`fixed_effect`, `index`, and `cluster`). 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_fenegbin <- dsl(model = "fenegbin", 
                    formula = Y ~ X1 + X2,
                    predicted_var = "Y",
                    prediction = "pred_Y",
                    fixed_effect = "twoways", 
                    index = c("unit", "year"), 
                    cluster = "unit",
                    data = data_fepois)
summary(out_fenegbin)
```

Fixed effects are handled as in `model = "fepois"`. The coefficients solve the moments of `fepois` weighted by the negative binomial weights `theta/(theta + mu)`, and `theta` is estimated by maximum likelihood using the labeled observations. In maximum likelihood (e.g., `fixest::fenegbin`), `mu` in the weights uses the estimated fixed effects. With DSL, the fixed effect of a group depends heavily on its few labeled observations (weighted by the inverse of the sampling probability), and weights that depend on them bias the coefficients (the incidental parameter problem). Thus, `fenegbin` computes `mu` in the weights with fixed effects estimated only from predictions. The moments are then valid for any weights, as in `model = "fepois"`, and the weights only affect efficiency.

