---
title: "Models Available in DSL"
output: 
  md_document:
    variant: markdown_github
---

<div class="card bg-light mb-3">
<div class="card-header">**Table of Contents**</div>
<div class="card-body">

- [**`lm`: Linear Regression**](#lm) <br>

- [**`logit`: Logistic Regression**](#logit) <br>

- [**`poisson`: Poisson Regression**](#poisson) <br>

- [**`felm`: Fixed Effects Linear Regression**](#felm) <br>

- [**`fepois`: Fixed Effects Poisson Regression**](#fepois) <br>


</div>
</div>

Note: We currently cover five models (`lm`, `logit`, `poisson`, `felm`, `fepois`). In the next update, we are planning to add more models, including multinomial-logit regression and two-stage least squares for the instrumental variable method. 

<br>

### **`lm`: Linear Regression**{#lm} 
<p class="lh-lg">
</p>

Researchers can implement DSL linear regression by specifying `model = "lm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
library(dsl)
data("data_lm") # example data 
head(data_lm)
```

In this example, variable `Y` requires text annotation (they take `NA` if a given observation is not labeled by experts). Variable `pred_Y` represents predictions for `Y`.

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_lm <- dsl(model = "lm", 
              formula = Y ~ X1 + X2 + X3 + X4 + X5,
              predicted_var = "Y",
              prediction = "pred_Y",
              data = data_lm)
```

| Argument | Description              |
|:-------- | :------------------------|
|`model` | A regression model. `dsl` currently supports `lm` (linear regression), `logit` (logistic regression), `poisson` (Poisson regression), `felm` (fixed-effects regression), and `fepois` (fixed-effects Poisson regression).|
|`formula` | A formula used in the specified regression model.|
|`predicted_var` | A vector of column names in the data that correspond to variables that need to be predicted.|
|`prediction` | A vector of column names in the data that correspond to predictions of `predicted_var`. |
| `data`    | A data frame. The class should be `data.frame`.|

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
summary(out_lm)
```

<br>

### **`logit`: Logistic Regression**{#logit} 
<p class="lh-lg">
</p>

Researchers can implement DSL logistic regression by specifying `model = "logit"`. It requires the same arguments as `model = "lm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_logit") # example data 
head(data_logit)
```

In this example, variables `Y` and `X1` require text annotation (they take `NA` if a given observation is not labeled by experts). Variables `pred_Y` and `pred_X1` represent predictions for `Y` and `X1`. As in this example, `dsl` can handle cases when multiple variables require text annotations.

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_logit <- dsl(model = "logit", 
                 formula = Y ~ X1 + X2 + X4,
                 predicted_var = c("Y", "X1"),
                 prediction = c("pred_Y", "pred_X1"),
                 data = data_logit)
summary(out_logit)
```

<br>

### **`poisson`: Poisson Regression**{#poisson} 
<p class="lh-lg">
</p>

Researchers can implement DSL Poisson regression for count outcomes by specifying `model = "poisson"`. It requires the same arguments as `model = "lm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_poisson") # example data 
head(data_poisson)
```

In this example, variable `Y` is a count (e.g., the number of times a topic is mentioned in a document) that requires text annotation (it takes `NA` if a given observation is not labeled by experts). Variable `pred_Y` represents predicted counts for `Y`.

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_poisson <- dsl(model = "poisson", 
                   formula = Y ~ X1 + X2 + X3 + X4,
                   predicted_var = "Y",
                   prediction = "pred_Y",
                   data = data_poisson)
summary(out_poisson)
```

As in a standard Poisson regression, coefficients are on the log scale, and `exp()` of a coefficient is the multiplicative change in the expected count (the incidence rate ratio). Because `dsl` estimates standard errors with the robust (sandwich) variance, `model = "poisson"` is Poisson Pseudo Maximum Likelihood (PPML). It only assumes that the expected count is `exp()` of the linear predictor, so the outcome does not need to follow a Poisson distribution (e.g., overdispersed counts are allowed) or be an integer. It should be non-negative.

<br>

### **`felm`: Fixed Effects Linear Regression**{#felm} 
<p class="lh-lg">
</p>

Researchers can implement DSL fixed effects linear regression by specifying `model = "felm"`. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_lm")  # example data 
head(data_felm)
```

In this example, variable `log_gsp` requires text annotation (it takes `NA` if a given observation is not labeled by experts). Variable `pred_log_gsp` represents predictions for `log_gsp`. 


#### One-way Fixed Effects

To implement DSL one-way fixed effects regression, users can set `fixed_effect = "oneway"`. Use `index` to denote which column defines fixed effects. Users can also cluster standard errors by specifying a variable name in argument `cluster` (this `cluster` argument is also avaiable for other models). 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_felm_one <- dsl(model = "felm", 
                    formula = log_pcap ~ log_gsp + log_pc + unemp,
                    predicted_var =  "log_gsp",
                    prediction = "pred_log_gsp",
                    fixed_effect = "oneway", 
                    index = c("state"), 
                    cluster = "state",
                    data = data_felm)
```
| Argument | Description              |
|:-------- | :------------------------|
|`fixed_effect` | A type of fixed effects regression you run. `oneway` (one-way fixed effects) or `twoways` (two-way fixed effects).|
|`index` | A vector of column names specifying fixed effects. When `fixed_effect = oneway`, it has one element. When `fixed_effect = twoways`, it has two elements, e.g., `index = c("state", "year")`. |
|`cluster` | A column name in the data that indicates the level at which cluster standard errors are calculated. Default is `NULL`.|


```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
summary(out_felm_one)
```

#### Two-way Fixed Effects

To implement DSL two-way fixed effects regression, users can set `fixed_effect = "twoways"`. Use `index` (a vector of length 2) to denote which columns define fixed effects. 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_felm_two <- dsl(model = "felm", 
                    formula = log_pcap ~ log_gsp + log_pc + unemp,
                    predicted_var =  "log_gsp",
                    prediction = "pred_log_gsp",
                    fixed_effect = "twoways", 
                    index = c("state", "year"), 
                    cluster = "state",
                    data = data_felm)
```

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
summary(out_felm_two)
```

<br>

### **`fepois`: Fixed Effects Poisson Regression**{#fepois} 
<p class="lh-lg">
</p>

Researchers can implement DSL fixed effects Poisson regression for count outcomes by specifying `model = "fepois"`. It uses the same arguments as `model = "felm"` (`fixed_effect`, `index`, and `cluster`). 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
data("data_fepois") # example data 
head(data_fepois)
```

In this example, variable `Y` is a count that requires text annotation (it takes `NA` if a given observation is not labeled by experts), and variable `pred_Y` represents predicted counts for `Y`. The data is a panel of 100 units over 20 years (`unit` and `year`). 

```{r eval = TRUE, echo = TRUE, tidy = FALSE, warning = FALSE, error = FALSE, message = FALSE}
out_fepois <- dsl(model = "fepois", 
                  formula = Y ~ X1 + X2,
                  predicted_var = "Y",
                  prediction = "pred_Y",
                  fixed_effect = "twoways", 
                  index = c("unit", "year"), 
                  cluster = "unit",
                  data = data_fepois)
summary(out_fepois)
```

As in `model = "poisson"`, coefficients are on the log scale and the estimator is Poisson Pseudo Maximum Likelihood (PPML), which allows overdispersed and non-integer outcomes. `fixed_effect = "oneway"` with one element in `index` implements one-way fixed effects.

We recommend `model = "fepois"` over adding fixed effects as dummy variables (e.g., `factor(unit)`) to `model = "poisson"`. The DSL estimator of the fixed effect of each group relies on the DSL-corrected sum of the outcome within the group. When a group is small, one labeled observation with a large prediction error can make this sum negative, and the fixed effect (the log of a negative number) is not defined. `model = "fepois"` concentrates out the fixed effects so that the coefficients are estimated even in such cases. When all fixed effects are defined, it gives the same estimates and standard errors as the dummy-variable approach. With `fixed_effect = "twoways"`, the fixed effects of the index with more levels are concentrated out, and those of the other index enter as dummy variables. 
